Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2023-05-12 doc/manuscript/ESSDmanuscript.Rmd
fix variable counts
partially fixes #46

C_variables <- read.csv(paste0(path_to_dir,"/doc/manuscript/figures_tables/C_variables.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_relevant_variables = nrow(C_variables) - 8 + 2  # n variables relevant = rows in C_variables - headers and total + 2 that are grouped with similar variables
n_relevant_variables_with_no_data =  sum(C_variables$`n in ForC` == 0,  na.rm = TRUE)
n_relevant_variables_with_data= n_relevant_variables - n_relevant_variables_with_no_data

@@

n_records_sent_to_EFDB = C_variables[length(C_variables[,"variable"]), "n sent to EFDB"] # n records sent to EFDB
n_relevant_variables_no_records_sent =  sum(C_variables$`n sent to EFDB` == 0,  na.rm = TRUE)
n_relevant_variables_records_sent= n_relevant_variables - n_relevant_variables_no_records_sent
n_records_sent_biomass<- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass"]
n_records_sent_agb <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_ag"]
n_records_sent_agb_components <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_ag_foliage"]
n_records_sent_rootbiomass <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_root"] 
n_records_sent_rootbiomass_components <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_root_coarse"]
n_records_sent_deadwood<- C_variables$`n sent to EFDB`[C_variables$`variable` == "deadwood"] 
n_records_sent_deadwood_components<- C_variables$`n sent to EFDB`[C_variables$`variable` == "deadwood_down"]+C_variables$`n sent to EFDB`[C_variables$`variable` == "deadwood_standing"]
n_records_sent_litter<- C_variables$`n sent to EFDB`[C_variables$`variable` == "litter"] +C_variables$`n sent to EFDB`[C_variables$`variable` == "O.horizon"]
n_records_sent_soil<- C_variables$`n sent to EFDB`[C_variables$`variable` == "SOM / SOC"] 
n_records_sent_stocks = n_records_sent_biomass+n_records_sent_agb+n_records_sent_agb_components+n_records_sent_rootbiomass+n_records_sent_rootbiomass_components+n_records_sent_deadwood+n_records_sent_deadwood_components+ n_records_sent_litter+ n_records_sent_soil
n_records_sent_delta.agb<- C_variables$`n sent to EFDB`[C_variables$`variable` == "delta.agb"]
n_records_sent_delta.litter<- C_variables$`n sent to EFDB`[C_variables$`variable` == "delta.O.horizon"]+C_variables$`n sent to EFDB`[C_variables$`variable` == "delta.litter"]
n_records_sent_increments <- n_records_sent_delta.agb+n_records_sent_delta.litter
n_records_sent_fluxes = n_records_sent_to_EFDB-n_records_sent_stocks-n_records_sent_increments

@@

We added a total of `r n_new_variables` new EFDB-relevant variables to the set of named and defined variables (Fig. `r fig_variable_mapping`), counting each pair of variables with units in C (ending in  `_C`) or organic matter (ending in `_OM`) as one. 
The majority of these were increment variables (n=11), adding to only one previously defined increment variable (aboveground biomass increment, *delta.agb*). 
These are directly related to C stocks as previously defined in ForC, with "*delta.*" added in front of the variable name.
Further, we added variables capturing the belowground component of woody mortality (*woody.mortality_root*) and the combined aboveground and belowground components of woody mortality (*woody.mortality*).
Although most of these variables lacked records in ForC as of `r format(Sys.time(), '%B %d, %Y')`, their addition gave the structure such that records can be populated over time. 
Finally, to provide better definition of the previously existing variable *organic.layer*, which has a nebulous definition that reflects the varied definitions adopted by original studies, we added two clearly defined variables: *litter* (relatively undecomposed plant material/ OL horizon), and *O.horizon* (entire O-horizon, including OL).

@@

ForC contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables relevant to EFDB (Table `r table_variables`, Fig. `r fig_variable_mapping`).
The records were very unevenly distributed across variables. 
The variable with most records was aboveground biomass, representing `r round(n_ind_records_agb/n_ind_records_EFDB_variables,2)*100`% of all independent records relevant to EFDB, and aboveground biomass components (woody biomass or foliage) representing an additional `r round(n_ind_records_agb_components/n_ind_records_EFDB_variables,2)*100`%.
A total of `r round(n_ind_records_rootbiomass/n_ind_records_EFDB_variables,2)*100`% of relevant records were for root biomass (including fine and coarse root components), while `r round(n_ind_records_biomass/n_ind_records_EFDB_variables,2)*100`% described total biomass.
The non-living pools were less represented, with `r round(n_ind_records_deadwood/n_ind_records_EFDB_variables,2)*100`% of relevant were for dead wood (including standing and fallen components), `r round(n_ind_records_OL/n_ind_records_EFDB_variables*100,1)`% for litter, `r round(n_ind_records_TotEcosystemC/n_ind_records_EFDB_variables*100,1)`% for total ecosystem C excluding soils, and `r round(n_ind_records_SOC/n_ind_records_EFDB_variables*100,1)`% for soil carbon.

@@

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table[,-1], format = "latex",  
             longtable = T,
             caption = "\\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.}") %>%
  row_spec(0, bold = TRUE)  %>%
  row_spec(which(C_variables_table$bold), bold = TRUE) %>% #total
  kable_styling(font_size = 8) %>%
  #kableExtra::kable_styling(full_width = TRUE) 
  kable_styling(latex_options = c("hold_position", "repeat_header"), protect_latex = T) 
```

@@

In terms of variables transferred, records were sent for `r n_relevant_variables_records_sent` variables, including variables from each C pool (Table `r table_variables`).
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records sent were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round(n_records_sent_rootbiomass/n_records_sent_to_EFDB,2)*100`% for root biomass, 
`r round(n_records_sent_rootbiomass_components/n_records_sent_to_EFDB,2)*100`% for components of root biomass (coarse or fine roots), 
`r round(n_records_sent_deadwood/n_records_sent_to_EFDB,2)*100`% for dead wood, 
`r round(n_records_sent_deadwood_components/n_records_sent_to_EFDB,2)*100`% for components of dead wood (standing or fallen), 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (OL layer or entire O horizon),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for SOM/ SOC.
Increment records totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records sent, including
`r round(n_records_sent_delta.agb/n_records_sent_to_EFDB*100,2)`% for aboveground biomass increments 
and `r round(n_records_sent_delta.litter/n_records_sent_to_EFDB*100,2)`% for organic horizon increments.
The remaining (`r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`%) of records sent described fluxes,
all of which were either inputs or outputs to the aboveground biomass pool,
and a subset of which also described inputs to the dead wood or litter pool (Table `r table_variables`, Fig. `r fig_variable_mapping`).

>>>>

C_variables <- read.csv(paste0(path_to_dir,"/doc/manuscript/figures_tables/C_variables.csv"), stringsAsFactors = FALSE, check.names = FALSE)
n_relevant_variables = nrow(C_variables) - 8  # n variables relevant = rows in C_variables - headers and total 
n_relevant_variables_with_no_data =  sum(C_variables$`n in ForC` == 0,  na.rm = TRUE)
n_relevant_variables_with_data= n_relevant_variables - n_relevant_variables_with_no_data

@@

n_records_sent_to_EFDB = C_variables[length(C_variables[,"variable"]), "n sent to EFDB"] # n records sent to EFDB
n_relevant_variables_no_records_sent =  sum(C_variables$`n sent to EFDB` == 0,  na.rm = TRUE)
n_relevant_variables_records_sent= n_relevant_variables - n_relevant_variables_no_records_sent
n_records_sent_biomass<- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass"]
n_records_sent_agb <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_ag"]
n_records_sent_agb_components <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_ag_foliage"]
n_records_sent_rootbiomass <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_root"] 
n_records_sent_rootbiomass_components <- C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n sent to EFDB`[C_variables$`variable` == "biomass_root_coarse"]
n_records_sent_deadwood<- C_variables$`n sent to EFDB`[C_variables$`variable` == "deadwood"] 
n_records_sent_deadwood_components<- C_variables$`n sent to EFDB`[C_variables$`variable` == "deadwood_down"]+C_variables$`n sent to EFDB`[C_variables$`variable` == "deadwood_standing"]
n_records_sent_litter<- C_variables$`n sent to EFDB`[C_variables$`variable` == "litter"] +C_variables$`n sent to EFDB`[C_variables$`variable` == "O.horizon"]
n_records_sent_soil<- C_variables$`n sent to EFDB`[C_variables$`variable` == "SOM / SOC"] 
n_records_sent_stocks = n_records_sent_biomass+n_records_sent_agb+n_records_sent_agb_components+n_records_sent_rootbiomass+n_records_sent_rootbiomass_components+n_records_sent_deadwood+n_records_sent_deadwood_components+ n_records_sent_litter+ n_records_sent_soil
n_records_sent_delta.agb<- C_variables$`n sent to EFDB`[C_variables$`variable` == "delta.agb"]
n_records_sent_delta.litter<- C_variables$`n sent to EFDB`[C_variables$`variable` == "delta.O.horizon"]+C_variables$`n sent to EFDB`[C_variables$`variable` == "delta.litter"]
n_records_sent_increments <- n_records_sent_delta.agb+n_records_sent_delta.litter
n_records_sent_fluxes <- sum(C_variables$`n sent to EFDB`[C_variables$`variable` %in% flux_variables_with_records])
n_records_sent_fluxes_check = n_records_sent_to_EFDB-n_records_sent_stocks-n_records_sent_increments

@@

To create structure for EFDB-relevant records, we added a total of `r n_new_variables` new variables to the set of named and defined variables (Fig. `r fig_variable_mapping`), counting each pair of variables with units in C (ending in  `_C`) or organic matter (ending in `_OM`) as one. 
The majority of these were increment variables (n=11), adding to only one previously defined increment variable (aboveground biomass increment, *delta.agb*). 
These are directly related to C stocks as previously defined in ForC, with "*delta.*" added in front of the variable name.
Further, we added variables capturing the belowground component of woody mortality (*woody.mortality_root*) and the combined aboveground and belowground components of woody mortality (*woody.mortality*).
Although most of these variables lacked records in ForC as of `r format(Sys.time(), '%B %d, %Y')`, their addition gave the structure such that records can be populated over time. 
Finally, to provide better definition of the previously existing variable *organic.layer*, which has a nebulous definition that reflects the varied definitions adopted by original studies, we added two clearly defined variables: *litter* (relatively undecomposed plant material/ OL horizon), and *O.horizon* (entire O-horizon, including OL).

@@

ForC contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables (or closely-related variable groups) relevant to EFDB (Table `r table_variables`, Fig. `r fig_variable_mapping`).
The records were very unevenly distributed across variables. 
The variable with most records was aboveground biomass, representing `r round(n_ind_records_agb/n_ind_records_EFDB_variables,2)*100`% of all independent records relevant to EFDB, and aboveground biomass components (woody biomass or foliage) representing an additional `r round(n_ind_records_agb_components/n_ind_records_EFDB_variables,2)*100`%.
A total of `r round(n_ind_records_rootbiomass/n_ind_records_EFDB_variables,2)*100`% of relevant records were for root biomass (including fine and coarse root components), while `r round(n_ind_records_biomass/n_ind_records_EFDB_variables,2)*100`% described total biomass.
The non-living pools were less represented, with `r round(n_ind_records_deadwood/n_ind_records_EFDB_variables,2)*100`% of relevant were for dead wood (including standing and fallen components), `r round(n_ind_records_OL/n_ind_records_EFDB_variables*100,1)`% for litter, `r round(n_ind_records_TotEcosystemC/n_ind_records_EFDB_variables*100,1)`% for total ecosystem C excluding soils, and `r round(n_ind_records_SOC/n_ind_records_EFDB_variables*100,1)`% for soil carbon.

@@

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table[,-1], format = "latex",  
             longtable = T,
             caption = "\\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.}") %>%
  row_spec(0, bold = TRUE)  %>%
  row_spec(which(C_variables_table$bold), bold = TRUE) %>% #total
  kable_styling(font_size = 8) %>%
  #kableExtra::kable_styling(full_width = TRUE) 
  kable_styling(latex_options = c("hold_position", "repeat_header"), protect_latex = T) 
```

@@

In terms of variables transferred, records were sent for `r n_relevant_variables_records_sent` variables (or closely-related variable groups), including variables from each C pool (Table `r table_variables`).
The majority (`r round(n_records_sent_stocks/n_records_sent_to_EFDB, 2)*100`%) of records sent were for C stocks, including
`r round(n_records_sent_biomass/n_records_sent_to_EFDB,2)*100`% for total biomass, 
`r round(n_records_sent_agb/n_records_sent_to_EFDB,2)*100`% for aboveground biomass,
`r round(n_records_sent_agb_components/n_records_sent_to_EFDB,2)*100`% for components of aboveground biomass (wood or foliage), 
`r round(n_records_sent_rootbiomass/n_records_sent_to_EFDB,2)*100`% for root biomass, 
`r round(n_records_sent_rootbiomass_components/n_records_sent_to_EFDB,2)*100`% for components of root biomass (coarse or fine roots), 
`r round(n_records_sent_deadwood/n_records_sent_to_EFDB,2)*100`% for dead wood, 
`r round(n_records_sent_deadwood_components/n_records_sent_to_EFDB,2)*100`% for components of dead wood (standing or fallen), 
`r round(n_records_sent_litter/n_records_sent_to_EFDB,2)*100`% for litter (entire O horizon or OL layer component),
and `r round(n_records_sent_soil/n_records_sent_to_EFDB,2)*100`% for SOM/ SOC.
Increment records totaled `r round(n_records_sent_increments/n_records_sent_to_EFDB*100, 0)`% of records sent, including
`r round(n_records_sent_delta.agb/n_records_sent_to_EFDB*100,2)`% for aboveground biomass increments 
and `r round(n_records_sent_delta.litter/n_records_sent_to_EFDB*100,2)`% for organic horizon increments.
The remaining (`r round(n_records_sent_fluxes/n_records_sent_to_EFDB*100, 0)`%) of records sent described fluxes,
all of which were either inputs or outputs to the aboveground biomass pool,
and a subset of which also described inputs to the dead wood or litter pool (Table `r table_variables`, Fig. `r fig_variable_mapping`).

### 2023-05-12 doc/manuscript/ESSDmanuscript.tex
fix variable counts
partially fixes #46

We added a total of 15 new EFDB-relevant variables to the set of named
and defined variables (Fig. 1), counting each pair of variables with
units in C (ending in \texttt{\_C}) or organic matter (ending in
\texttt{\_OM}) as one. The majority of these were increment variables
(n=11), adding to only one previously defined increment variable
(aboveground biomass increment, \emph{delta.agb}). These are directly
related to C stocks as previously defined in ForC, with
``\emph{delta.}'' added in front of the variable name. Further, we added
variables capturing the belowground component of woody mortality
(\emph{woody.mortality\_root}) and the combined aboveground and
belowground components of woody mortality (\emph{woody.mortality}).
Although most of these variables lacked records in ForC as of May 12,
2023, their addition gave the structure such that records can be
populated over time. Finally, to provide better definition of the
previously existing variable \emph{organic.layer}, which has a nebulous
definition that reflects the varied definitions adopted by original
studies, we added two clearly defined variables: \emph{litter}
(relatively undecomposed plant material/ OL horizon), and
\emph{O.horizon} (entire O-horizon, including OL).

@@

As of May 12, 2023, ForC (v4.0) contained 32693 independent records
(39855 total), 19323 of which were for the the 45 variables relevant to
EFDB (Fig. 1). These records were distributed across all forested
continents and ecozones, albeit unevenly (Fig. 2). The largest number of
records came from Asia, followed by North America, South America, and
Europe, with relatively few records from Africa, Australia, and Oceania
(Fig. 3c). Categorized by FAO ecozone, the greatest numbers of records
came from subtropical humid forests, temperate mountain systems, and
tropical rain forests, each with \textgreater2,000 independent records
(Fig. 3b). Boreal coniferous forests, temperate continental forests,
subtropical mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

ForC contained records for 31 of the 45 variables relevant to EFDB
(Table 2, Fig. 1). The records were very unevenly distributed across
variables. The variable with most records was aboveground biomass,
representing 42\% of all independent records relevant to EFDB, and
aboveground biomass components (woody biomass or foliage) representing
an additional 5\%. A total of 27\% of relevant records were for root
biomass (including fine and coarse root components), while 4\% described
total biomass. The non-living pools were less represented, with 4\% of
relevant were for dead wood (including standing and fallen components),
0.4\% for litter, 0.3\% for total ecosystem C excluding soils, and 2.1\%
for soil carbon.

@@

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1094 & 850 & 95 & 50\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9456 & 8155 & 1348 & 764\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 520 & 73 & 45\\
\hline
delta.agb & 166 & 150 & 145 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_woody\_stem & 949 & 622 & 60 & 61\\
\hline
ANPP\_woody\_branch & 243 & 200 & 4 & 4\\
\hline
woody.mortality\_ag & 112 & 75 & 47 & 50\\
\hline
stem\_pC & 9 & 0 & 0 & 0\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4185 & 125 & 57\\
\hline
biomass\_root\_fine & 930 & 595 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 413 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root & 577 & 416 & 0 & 0\\
\hline
BNPP\_root\_fine & 488 & 331 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 304 & 104 & 70\\
\hline
deadwood\_standing & 153 & 121 & 18 & 17\\
\hline
deadwood\_down & 425 & 369 & 52 & 28\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 45 & 45 & 45 & 40\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
ANPP\_litterfall & 294 & 253 & 11 & 11\\
\hline
NPP\_litter & 94 & 70 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Total Ecosystem C (excl. soils)} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
total.ecosystem\_2 & 64 & 64 & 0 & 0\\
\hline
delta.total.ecosystem\_2 & 0 & 0 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 401 & 89 & 56\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{23575} & \textbf{19323} & \textbf{2283} & \textbf{1438}\\
\hline
\end{longtable}
\endgroup{}

@@

In terms of variables transferred, records were sent for 21 variables,
including variables from each C pool (Table 2). The majority (82\%) of
records sent were for C stocks, including 3\% for total biomass, 53\%
for aboveground biomass, 4\% for components of aboveground biomass (wood
or foliage), 4\% for root biomass, 2\% for components of root biomass
(coarse or fine roots), 5\% for dead wood, 3\% for components of dead
wood (standing or fallen), 4\% for litter (OL layer or entire O
horizon), and 4\% for SOM/ SOC. Increment records totaled 9\% of records
sent, including 8.55\% for aboveground biomass increments and 0.28\% for
organic horizon increments. The remaining (9\%) of records sent
described fluxes, all of which were either inputs or outputs to the
aboveground biomass pool, and a subset of which also described inputs to
the dead wood or litter pool (Table 2, Fig. 1).

>>>>

To create structure for EFDB-relevant records, we added a total of 15
new variables to the set of named and defined variables (Fig. 1),
counting each pair of variables with units in C (ending in \texttt{\_C})
or organic matter (ending in \texttt{\_OM}) as one. The majority of
these were increment variables (n=11), adding to only one previously
defined increment variable (aboveground biomass increment,
\emph{delta.agb}). These are directly related to C stocks as previously
defined in ForC, with ``\emph{delta.}'' added in front of the variable
name. Further, we added variables capturing the belowground component of
woody mortality (\emph{woody.mortality\_root}) and the combined
aboveground and belowground components of woody mortality
(\emph{woody.mortality}). Although most of these variables lacked
records in ForC as of May 12, 2023, their addition gave the structure
such that records can be populated over time. Finally, to provide better
definition of the previously existing variable \emph{organic.layer},
which has a nebulous definition that reflects the varied definitions
adopted by original studies, we added two clearly defined variables:
\emph{litter} (relatively undecomposed plant material/ OL horizon), and
\emph{O.horizon} (entire O-horizon, including OL).

@@

As of May 12, 2023, ForC (v4.0) contained 32693 independent records
(39855 total), 19323 of which were for the the 43 variables relevant to
EFDB (Fig. 1). These records were distributed across all forested
continents and ecozones, albeit unevenly (Fig. 2). The largest number of
records came from Asia, followed by North America, South America, and
Europe, with relatively few records from Africa, Australia, and Oceania
(Fig. 3c). Categorized by FAO ecozone, the greatest numbers of records
came from subtropical humid forests, temperate mountain systems, and
tropical rain forests, each with \textgreater2,000 independent records
(Fig. 3b). Boreal coniferous forests, temperate continental forests,
subtropical mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

ForC contained records for 29 of the 43 variables (or closely-related
variable groups) relevant to EFDB (Table 2, Fig. 1). The records were
very unevenly distributed across variables. The variable with most
records was aboveground biomass, representing 42\% of all independent
records relevant to EFDB, and aboveground biomass components (woody
biomass or foliage) representing an additional 5\%. A total of 27\% of
relevant records were for root biomass (including fine and coarse root
components), while 4\% described total biomass. The non-living pools
were less represented, with 4\% of relevant were for dead wood
(including standing and fallen components), 0.4\% for litter, 0.3\% for
total ecosystem C excluding soils, and 2.1\% for soil carbon.

@@

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.} \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1094 & 850 & 95 & 50\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9456 & 8155 & 1348 & 764\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 520 & 73 & 45\\
\hline
delta.agb & 166 & 150 & 145 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_woody\_stem & 949 & 622 & 60 & 61\\
\hline
ANPP\_woody\_branch & 243 & 200 & 4 & 4\\
\hline
woody.mortality\_ag & 112 & 75 & 47 & 50\\
\hline
stem\_pC & 9 & 0 & 0 & 0\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4185 & 125 & 57\\
\hline
biomass\_root\_fine & 930 & 595 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 413 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root & 577 & 416 & 0 & 0\\
\hline
BNPP\_root\_fine & 488 & 331 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 304 & 104 & 70\\
\hline
deadwood\_standing & 153 & 121 & 18 & 17\\
\hline
deadwood\_down & 425 & 369 & 52 & 28\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 45 & 45 & 45 & 40\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
ANPP\_litterfall & 294 & 253 & 11 & 11\\
\hline
NPP\_litter & 94 & 70 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Total Ecosystem C (excl. soils)} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
total.ecosystem\_2 & 64 & 64 & 0 & 0\\
\hline
delta.total.ecosystem\_2 & 0 & 0 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 401 & 89 & 56\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{23575} & \textbf{19323} & \textbf{2283} & \textbf{1438}\\
\hline
\end{longtable}
\endgroup{}

@@

In terms of variables transferred, records were sent for 19 variables
(or closely-related variable groups), including variables from each C
pool (Table 2). The majority (82\%) of records sent were for C stocks,
including 3\% for total biomass, 53\% for aboveground biomass, 4\% for
components of aboveground biomass (wood or foliage), 4\% for root
biomass, 2\% for components of root biomass (coarse or fine roots), 5\%
for dead wood, 3\% for components of dead wood (standing or fallen), 4\%
for litter (entire O horizon or OL layer component), and 4\% for SOM/
SOC. Increment records totaled 9\% of records sent, including 8.55\% for
aboveground biomass increments and 0.28\% for organic horizon
increments. The remaining (9\%) of records sent described fluxes, all of
which were either inputs or outputs to the aboveground biomass pool, and
a subset of which also described inputs to the dead wood or litter pool
(Table 2, Fig. 1).
